---
title: "Weather Data Report"
author: "Phil Alderman, Ryan Earp, and Katherine Haile"
date: "12/2/2021"
output: html_document
---
<<<<<<< HEAD
   
# Summary   
This, that, and the other thing 
   
=======

## Data Consolidation

  The first task to combine the excel sheets created by the library staff from the paper documents was achieved using an function we created in R. The filename of the document was used as the argument of the function, and the read_excel and read_ods functions were used to assign the file to a temporary dataframe called raw. The raw dataframes were then subsetted to exclude the summary data and column names used in the original documents and clean variable names were assigned to the columns. Due to the variation in characters in some of the numerical columns, these were coerced to a numeric form and the original columns were labeled as such. Columns to record the year, month, observer were assigned to the corresponding cells in the original document. The overall date was also combined and the times for the beginning and ending time of rain events were set as such. Finally, the date was set as the first column and the year, month, and day columns were removed.
  This function was then used to read in all of the original files and create a single .csv file. The file names were extracted from the raw data folder and ensured only the correct files were selected. This list was then fed into the read_wth_data function and arranged by date.

```{r}
read_wth_data <- function(file_name){
  if(str_detect(file_name, "1937-11")){
    sheet = "Nov. 1937"
  }else{
    sheet = 1
  }
  if(str_detect(file_name, "1896-01")){
    cshift <- 1
  }else{
    cshift <- 0
  }
  if(str_detect(file_name, "\\.ods$")){
    raw <- read_ods(file_name, col_names = FALSE, sheet = sheet) %>% 
      as_tibble() %>% 
      mutate(across(where(is.numeric), .fns = as.character))
  }else{
    raw <- read_excel(file_name, col_names = FALSE, sheet = sheet)
  }
  wdata <- raw[30:60, 1:14] %>% 
    rename_with(.fn = ~c("day", "temp_maximum", "temp_minimum", "temp_range", "temp_set_max", 
                    "precip_time_of_beginning", "precip_time_of_ending", "precip_amount",
                    "precip_snowfall_in_inches", "precip_snow_depth_tobs",
                    "wind_dir_tobs", "weather_state_tobs", "wind_dir_day", "weather_state_day")) %>% 
    mutate(across(c(temp_maximum, temp_minimum, temp_range, temp_set_max,
                    precip_amount, precip_snowfall_in_inches,
                    precip_snow_depth_tobs),
                  list(num = as.numeric))) %>% 
    rename_with(c(temp_maximum, temp_minimum, temp_range, temp_set_max,
                  precip_amount, precip_snowfall_in_inches,
                  precip_snow_depth_tobs,
                  precip_time_of_beginning, precip_time_of_ending),
                .fn = ~str_c(.,"_orig")) %>% 
    rename_with(matches("_num$"),
                .fn = ~str_remove(.,"_num$")) %>% 
    mutate(year = sanitize_year(as.character(raw[3, 4 + cshift])),
           month = sanitize_month(as.character(raw[3, 2 + cshift])),
           observer = as.character(raw[64, 2]),
           date = mdy(paste(month, day, year)),
           precip_time_of_beginning = format(times(as.numeric(precip_time_of_beginning_orig))),
           precip_time_of_ending = ifelse(is.na(as.numeric(precip_time_of_ending_orig)),
                                          precip_time_of_ending_orig,
                                          format(times(as.numeric(precip_time_of_ending_orig))))) %>% 
    select(date, everything(), -c(day, month, year))
    
return(wdata)
}
source("code/read_weather_data.R")
wth_data <- list.files("data/raw/Weather Record",
                        "^[0-9]{4}$",
                        full.names = TRUE) %>% 
   map(~list.files(., "(\\.xlsx)|(\\.ods)", full.names = TRUE)) %>% 
   unlist() %>% 
   str_subset("(1893_04\\.xlsx)|(~\\$)", negate = TRUE) %>% 
   tibble(file_path = .) %>% 
   mutate(imported_data = map(file_path, ~try(read_wth_data(.)))) %>% 
   unnest(imported_data) %>% 
   arrange(date)

wth_data %>%
   filter(!is.na(date)) %>% 
   arrange(date) %>% 
   write_csv("data/processed/imported_weather_data_1893-1940.csv")
```

## Fixed Errors in Import Process

  During the import process, some challenges were encountered in reading in the files. Two different types of files were used to document the written documents, excel and ods. Due to this, two different functions were needed to read in the files to R. Additionally, varying naming conventions for the files were used throughout the years with some duplicated, so the file list needed to be subsetted to only include the desired files.
  Additionally, some errors were reported when trying to read in some of the files. One file had the sheet name labeled as the month and year instead of sheet 1 like the rest, so an exception was needed for this file. Another file had the summary titles in column A repeated in column B so the information collected from this needed to be shifted to the right one. 
  After an initial summary of the data set, it was observed that there were several entries were recorded after the year 2000 and others the date was NA. Further investigation found errors that were corrected in code. One file contained the incorrect year in the year cell. Others had varying errors in the month column of the document. This occured from either mispelling of the month or using a mispelled abbreviation caused the function to incorrectly assign the year.
```{r}
sanitize_year <- function(year){
  if(year == "2021"){
    year <- "1921"
  }
  return(year)
}

sanitize_month <- function(month){
  if(month %in% c("Fabruary", "Feburary")){
    month <- "February"
  }else if(month %in% c("Semptember", "Spet.")){
    month <- "September"
  }
  return(month)
}
```
>>>>>>> 2c7b8060132d45728861c0a51796ac20844c86aa
                         
# Wind Direction at Time of Observation (wind_dir_tobs)  
 
                 

```{r wind_dir_tobs}



```


# Wind Direction at Time of Observation (wind_dir_tobs)  
  To check for any errors within the wind direction column, the unique function was used to show all unique entries within the column. This resulted in a list that contained variations on how the directions were inputted as capitalization changes would result in a unique entry. Most of these were able to be resolved using code as we could input all the variations and spelling errors and say what it needed to be. This leaves four data entries where the entry is uncertain and will need to be checked.

```{r wind_dir_tobs}
unique_wind_dir <- unique(wth_data$wind_dir_tobs)

sanitize_wind_dir <- function(wind_dir){
  clean_wind_dir <- str_replace_all(wind_dir, c("( *\\& *)|( *– *)|( *\\+ *)|(;(?=[^ ]))" = "; ",
                                                "Wind direction at time of observation" = "W",
                                                "East" = "E",
                                                "South" = "S",
                                                "Sw" = "SW",
                                                "SS" = "SW",
                                                "ES" = "SE",
                                                "Souh" = "S",
                                                "NR" = "NE"))
wth_data %>% filter(wind_dir_tobs == "?" | wind_dir_tobs == "NE; N?")

```


# Weather State at Time of Observation (weather_state_tobs)
  Similarly to the the wind direction, there are several errors within the weather state due to either misspelling or different ways of entering the same data entry. A similar code will need to be developed for the spelling errors identified within in this column. There are seven data entries containing question marks indicating the value is uncertain and will need to be manually checked.


```{r weather_state_tobs}
unique_weather <- unique(wth_data$weather_state_tobs) 

wth_data %>% 
filter(weather_state_tobs == "Pt????" | weather_state_tobs == "?")
```

# Wind Direction for the Day (wind_dir_day)



```{r wind_dir_day}



```


# Weather State for the Day (weather_state_day)
There are 49 factors of this variable in the final data set. The original instructions call for only using cloudy, partly cloudy, and clear for the state of the weather
```{r weather_state_day}
wth_data <- read_csv("data/processed\\imported_weather_data_1893-1940.csv")
bad_states <- wth_data %>% 
  select(weather_state_day) %>%
  unique() 

wth_data %>% 
  group_by(weather_state_day) %>% 
  ggplot(aes(x = weather_state_day)) +
  geom_bar()
```


# Maximum and Minimum Temperature (temp_maximum and temp_minimum)
  When evaluating the minimum and maximum temperature the data frame was first checked for NA values, which resulted 191 observations. These were checked against the original documents, and the majority were the result of missing values in the original data set. However, the months of May through August in 1929 were not recorded at all and will need to be filled in.
Additionally, the data frame was check to see if any observations had recorded values where the minimum temperature for the day was greater than the maximum temperature for the day. There are 27 observations that fit this, and after some further evaluation, some are the result of errors in the original data set and others from misreading the documents. It is recommended that these are examined individually to determine the proper action needed.
```{r temp}
na_temp <- wth_data %>% 
  filter(is.na(temp_maximum), is.na(temp_minimum))
```



# Temperature Range for the Day (temp_range)
  To check the temperature range column, the dataframe was filtered to include any observations where the range did not equal the maximum minus the minimum temperature. A total of 4340 observations fit this description. No further cleaning was done for this as the error could occur within any of the three columns.
```{r temp_range}
wrong_range <- wth_data %>% 
  filter(temp_range != temp_maximum - temp_minimum) %>% 
  mutate(diff =(temp_maximum - temp_minimum)-temp_range) %>% 
  select(file_path, date, temp_minimum, temp_maximum, temp_range, diff)

wrong_range %>% 
  ggplot(aes(x = diff))+
  geom_histogram()
```
# Precipitation Amount (precip_amount)
There are six values found that were likely transcribed incorrectly from the original data sheet. Four of the highest values simply need a decimal placed before the first number to make them correct. The fifth and sixth incorrect values, which are the next highest ones in ranking order, are difficult to interpret from the original handwriting. However, these values are not correct when compared to the monthly totals that are listed. It may also be worthwhile to import the monthly totals entered in the data sheets and compare them to the calculated totals from each entry for each respective month, which would help ensure that there are no other mistyped data entries that are not as easily noticeable.
```{r precip_amount}
wth_data <- read_csv("data/processed\\imported_weather_data_1893-1940.csv")

wth_data %>% 
  ggplot()+
  geom_line(aes(x = date, y = precip_amount)) 

precip <- wth_data %>% 
  select(date, precip_amount) %>% 
  drop_na() %>% 
  arrange(desc(precip_amount))
```


# Snowfall in Inches (precip_snowfall_in_inches)
```{r precip_snowfall_in_inches}

```


# Snow Depth at Time of Observation (precip_snow_depth_tobs)
```{r precip_snow_depth_tobs}

```


# Beginning Time of Precipitation (precip_time_of_beginning)
```{r precip_time_of_beginning}

```


# Ending Time of Precipitation (precip_time_of_ending)
```{r precip_time_of_ending}

```





